---
title: "Requirements Quality Impact on Missing Entities"
author: "Julian Frattini"
date: '2023-10-10'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("../util/setup.R")
source("../util/model-eval.R")
```

This document contains the investigation of the impact of **requirements quality** on the number of **missing entities** in the process of domain modeling .

## Data Loading

```{r data}
source("../util/data-preprocessing.R")
d <- load.data()
```

## Bayesian Data Analysis 

We perform a Bayesian data analysis to answer the question: which factors influence the number of missing entities from a domain model.

### Graphical Causal Model

We draw a Directed Acyclic Graph (DAG) to make our causal assumptions explicit.

```{r dag}
dag <- dagify(
  me ~ rq + dur + edu + exp + dom + task + tool,
  dom ~ exp,
  exposure = "rq",
  outcome = "me",
  labels = c(me = "missing.entities", dur = "duration",
             rq = "requirements.quality",
             edu = "education", exp="experience",
             dom = "domain.knowledge", task = "task.experience",
             tool = "tool.experience"),
  coords = list(x=c(rq=1, me=2, dur=1.5, edu=1, exp=1, dom=1.2, task=1, tool=1),
                y=c(rq=-1, me=-1.5, dur=0, edu=-2, exp=-3.5, dom=-4, task=-2.5, tool=-3))
)

ggdag_status(dag, 
             use_labels = "label", 
             text = FALSE) +
  guides(fill = "none", color = "none") + 
  theme_dag()
```

We do not assume a relationship between *requirements quality* and *duration* as we have already disproven this relationship in the [duration](./duration.Rmd) analysis.

### Formula Definition and Prior Selection

We derive our formula from our DAG to infer the strength of causal relationships from it.

#### Model Definition

Based on the DAG, we construct a model with the following parameters:

| Term | Predictor | Rationale |
|---|---|---|
| `1` | Intercept | General challenge of missing an entity |
| `(1|PID)` | Participant-specific variability | Participant-specific defiation from the intercept in generally missing an entity |
| `RQ*period` | Carryover | Interaction effect between the treatment and the period, i.e., whether some treatments have a stronger effect in later periods |
| (`RQ`) |  Main Factor | Influence of a requirements quality defect on connecting entities wrongly |
| (`period`) | Sequence effect | Period in which the subject applies the treatment |
| `rel.duration` | Time | (Self-made) time pressure to complete the domain model |
| *rest* | Confounders | Context factors influencing the causal relationships. |

Predictors in brackets (e.g., (`RQ`)) are implied by the interaction effects (e.g., `RQ*dom.os`) and do not explicitly appear in the formula.

```{r formula}
formula <- entities.missing | trials(entities.expected) ~ 1 + 
  (1|PID) + 
  RQ*period +
  rel.duration + 
  exp.se.scaled + exp.re.scaled + edu + primary.role +
  dom.os + dom.db + 
  model.train + model.occ
```

For each predictor, we select an appropriate prior that reflects our prior belief about the causal relationship. We use weakly informative priors to allow the model in the later estimation step to infer the strength of the causal relations itself.

```{r priors}
priors <- c(
  prior(normal(-1, 0.2), class=Intercept),
  prior(normal(0, 0.1), class=b),
  prior(weibull(2, 1), class=sd)
)
```

#### Prior Evaluation

We sample directly from the priors (i.e., sample without the model updating its parameter distributions based on the observed data) to evaluate the feasibility of the priors.

```{r model-sample-priors}
m.prior <-
  brm(data = d, family = binomial, formula, prior = priors,
    iter = 4000, warmup = 1000, chains = 4, cores = 4,
    seed = 4, sample_prior="only",
    file = "fits/missing.entities.prior"
  )
```

With the samples from the prior distributions, we conduct a prior predictive check to ensure that the priors are appropriate, i.e., the actual observations are realistic given the prior knowledge of the model.

```{r prior-predictive-check}
ndraws <- 100
priorpc <- brms::pp_check(type="bars", m.prior, ndraws=ndraws)
priorpc
```

The distribution of draws encompasses the actually observed data, meaning that the actual observations lie within the realm of belief defined by the priors. We accept our priors as feasible.

### Model Training

Now, we execute the estimation step by updating the prior distributions based on the observed data `d`.

```{r model}
m <-
  brm(data = d, family = binomial, formula, prior = priors,
    iter = 4000, warmup = 1000, chains = 4, cores = 4,
    seed = 4, 
    file = "fits/missing.entities"
  )
```

We perform a posterior predictive check to ensure that the model has learned properly.

```{r posterior-predictive-check}
brms::pp_check(m, type="bars", ndraws=ndraws)
```

The distribution of draws still encompasses the actually obsered data. Additionally, the distribution grew narrower around the observed data, indicating that the posterior distributions more accurately reflect the the observed data.

Because we only trained one model, there is no need to select the best performing model using leave-one-out comparison `loo_compare`.

### Comparison regarding conditional Independencies

We test for conditional independencies. According to our DAG, we assume that $exp.se \rightarrow missing.entities$, $exp.se \rightarrow dom$, and $dom \rightarrow missing.entities$. We test whether the first and/or last hypothesis are true by comparing the selected model with other candidates where one of the two factors is missing. 

```{r comparison-formula}
f.no.exp <- entities.missing | trials(entities.expected) ~ 1 + 
  (1|PID) + 
  RQ*period +
  rel.duration + 
  exp.re.scaled + edu + primary.role +
  dom.os + dom.db + 
  model.train + model.occ

f.no.dom <- entities.missing | trials(entities.expected) ~ 1 + 
  (1|PID) + 
  RQ*period +
  rel.duration + 
  exp.se.scaled + exp.re.scaled + edu + primary.role +
  model.train + model.occ
```

```{r model-skew-comp}
m.no.exp <-
  brm(data = d, family = binomial, f.no.exp, prior = priors,
    iter = 4000, warmup = 1000, chains = 4, cores = 4,
    seed = 4, 
  )

m.no.dom <-
  brm(data = d, family = binomial, f.no.dom, prior = priors,
    iter = 4000, warmup = 1000, chains = 4, cores = 4,
    seed = 4
  )
```

We compare the predictors of interest ($\beta_{exp.se\_scaled}$ and $\beta_{dom.os}$) to see whether the estimates change when including either of the variables.

```{r conditional-independencies}
param.m <- mcmc_intervals_data(m)
param.m.no.exp <- mcmc_intervals_data(m.no.exp)
param.m.no.dom <- mcmc_intervals_data(m.no.dom)

param.m$model <- "full model"
param.m.no.exp$model <- "no exp.se"
param.m.no.dom$model <- "no dom"

combined.estimates <- rbind(param.m, param.m.no.exp, param.m.no.dom) %>% 
  filter(parameter %in% c("b_exp.se.scaled", 
                          "b_dom.os.L", "b_dom.os.Q", "b_dom.os.C", "b_dom.os.E4", 
                          "b_dom.db.L", "b_dom.db.Q", "b_dom.db.C"))

position.offset <- position_nudge(
  y = ifelse(
    combined.estimates$model == "full model",
    -0.1,
    ifelse(combined.estimates$model == "no dom", 0.1, 0)))

ggplot(data = combined.estimates,
       mapping = aes(x = m, y = parameter, color = model)) +
  geom_linerange(aes(xmin = l, xmax = h), position = position.offset, linewidth = 2) +
  geom_linerange(aes(xmin = ll, xmax = hh), position = position.offset) +
  geom_point(position = position.offset, color="black") + 
  geom_vline(xintercept = 0, linetype="dashed")
```

It becomes apparent that none of the variables are conditionally independent of each other: learning either variable does not diminish the information gained from the other variables. Consequently, we assume all three causal assumptions being correct.

### Model selection

We visualize the summary of updated posterior distributions - which does not yet answer the initial question, but indicates the direction and strength that each predictor has on the response variable.

```{r model-summary}
summary(m)
```

### Model Evaluation

#### Posterior Predictions

We let the model predict the response variable with a fixed independent variable (RQ=1 (passive voice), RQ=2 (ambiguous pronoun), and RQ=3 (passive voice + ambiguous pronoun)) and representative values for the context factors. This posterior prediction takes all uncertainty, which the model picked up, into account.

```{r posterior-comparison-passive}
evaluate.model(model=m, treatment=1)
```

```{r posterior-comparison-pronoun}
evaluate.model(model=m, treatment=2)
```

```{r posterior-comparison-passive-pronoun}
evaluate.model(model=m, treatment=3)
```

The posterior predictions show a slight effect of the use of ambiguous pronouns on the number of missing entities.

These values can also be obtained via `brms::conditional_effects`: by selecting the treatment `RQ` as the effect, the `posterior_epred` method predicts the distribution of the expected value when fixing the treatment to one of the four 

```{r posterior-comparison}
conditional_effects(m, effects="RQ", method="posterior_epred")
```

#### Marginal and Interaction Effects

We visualize the interaction effect between the period variable (i.e., the time slot in which a specific requirements specification was processed) and the treatment to identify any potential carryover effects.

```{r interaction-treatment-period}
conditional_effects(m, effects=c("period", "period:RQ"))
```

The visualizations show a slight overall learning effect (i.e., a slight negative slope of the `period` variable) which is diversified, but fairly negligible for the different treatments. 
